Road Segment Prioritization for Bicycle Infrastructure

### 2021-04-01 README.Rmd
Update README.Rmd

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. This paper demonstrates a new approach to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. Although each of these methods have been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach. The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios. A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  bookdown::pdf_document2:
  # # bookdown::word_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
editor_options: 
  markdown: 
    wrap: 72
---

@@

The 2015 Paris agreement [@agreement2015paris] acknowledged that
fundamental changes to societies and economies are necessary to mitigate
climate change. Like other sectors, transport is under substantial
pressure to decarbonise resulting in a number of technical innovations
including electric vehicles. But new vehicle technologies can only go so
far and do not tackle parallel problems such as congestion, road traffic
casualties and physical inactivity [@brand_climate_2020].

@@

> Inadequate cycling infrastructure discourages cycling and wastes
 public money. Much cycling infrastructure in this country is
 inadequate. It reflects a belief, conscious or otherwise, that hardly
 anyone cycles, that cycling is unimportant and that cycles must take
 no meaningful space from more important road users, such as motor
 vehicles and pedestrians [@departmentgearchange2020]

@@

Much research has been done to understand what gets people to cycle.
Segregated cycling infrastructure^[*Segregated cycling infrastructure*
refers to road space that is allocated to cyclists only, with physical
separation to protect cyclists from other modes of transport.] has been
shown to increase cycling uptake [@aldred2019impacts; @goodman2014new;
@marques2015infrastructure], with the separation from motorized
vehicles being key [@winters2011motivators]. Revealed preference of
cyclists shows that they are willing to deviate from the most efficient
routes in order to commute on safer roads [@crane2017longitudinal].
However, such deviations are only considered if they do not considerably
increase route circuitry; behaviour studies have found that the
probability of choosing a route decreases in proportion to its length
relative to the shortest route [@broach2011bicycle; @winters2010far].
Another defining feature for cycling infrastructure is how well
connected it is. Cyclists prefer cohesive infrastructure, particularly
when cycling on arterial roads with high levels of motorized traffic
[@stinson2003commuter]. The lack of well-connected cycling
infrastructure is one of the main obstacles to increasing cycling uptake
[@caulfield2012determining]. While direct and cohesive cycling networks
have been shown to positively impact cycling rates, density^[making an
area's bicycle network denser means adding more cycling routes in the
area and thereby giving cyclists more route options] of the cycling
network is also vital [@schoner2014missing].

@@

The studies outlined above lay out the fundamentals for designing
cycling networks that generate significant cycling uptake, but they do
not propose network-level interventions. In this section we outline
methods used in past studies, namely optimization and network
analysis methods, such as connected components and community
detection, and examine how they are leveraged to suggest cycling network
designs. We compare the effectiveness of these network-level studies in
incorporating the fundamentals outlined above. 
Our proposed approach is inspired by these methods, but it attempts to add to them
by ensuring that all of the outlined fundamentals are accounted for. It
also goes further by attempting to factor in ethical considerations
relating to distribution of investment.

*Optimization* methods have been used to propose improvements to
cycling networks. @mesbah2012bilevel propose a bi-level formulation to
optimize allocation of cycling lanes to the network without exceeding a
set budget. They account for the effect of cycling lanes on car traffic,
and attempt to maximize utilization of said lanes with minimal impact on
car travel times. Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as "traffic evaporation" [@nello2020environmental].
Car usage is not considered by @mauttone2017bicycle,
who develop an optimization framework that aims to minimize the total
user cost of cycling on the network. The aggregate flow^[*flow* is
used throughout this research to refer to the cycling demand when it is
routed onto the road network. The flow on any road segment is the
cumulative demand on it, resulting from cyclists commuting between
various OD pairs] on links^[*link*
refers to a road segment throughout this research] is obtained by using
shortest paths to route existing cycling demand onto the road network,
and the solution is a proposed set of links where cycling infrastructure
should be added in order to minimize the overall travel cost of cyclists
across the network. The cost of traversing a link is given as a function
of its length and whether or not it has cycling infrastructure. The
problem has also been solved by attempting to find the minimum cost of
improving roadway links to meet a desired level of service (LOS)
[@duthie2014optimization]. In this formulation, all origin-destination
(OD) pairs need to be connected by roads that meet the desired LOS. A
limitation of these problem formulations is that they do not explicitly solve for
continuity, which is dealt with using either a constraint specifying
that each link with a bike lane should be connected to at least one
destination [@mesbah2012bilevel], a constraint on maximum deviation
from shortest paths [@duthie2014optimization], or a discontinuity
penalty to prioritize connected road segments [@mauttone2017bicycle].

In this paper, continuity is analysed by looking at the connectivity of
the network through the graph-theoretic concept of *connected
components*. Other academics have taken this approach; @natera2019data
study the existing cycling network in terms of its disconnected
components and introduce two different algorithms to connect these
components by their most critical links and, in doing so, measure the
size of the growth of the largest connected component as a function of
the kilometers of network added. The concept of connected components is
also at the core of the methodology proposed by @olmos2020data. After
routing the cycling demand onto the network links, they use percolation
theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
filtering to identify the minimum flow at which the whole city is
connected by a giant component. While these approaches deal with
continuity better, they look at the network as a whole when attempting
to improve it, and in doing so fail to account for equitable
distribution of infrastructure.

@@

Our work builds on these *community finding* approaches by proposing a
similar greedy network expansion algorithm for cycle network expansion
within communities. We incorporate community finding methods for study
area partitioning with weighted routing to avoid links that are
stressful to cycle on. In doing so, we propose an approach that
accounts for motivators and deterrents to cycling. We propose three
sub-methods that address on some of the limitations of previous
studies. These limitations include (a) bias inherent when basing network
design solely on existing cycling demand, (b) proposing routes that may
not correspond to studies on cyclist preference and government policies,
and (3) insufficient consideration of the ethical principles underlying
the analysis. Section \@ref(calculating-potential-cycling-demand)
focuses on calculating potential cycling demand. Section \@ref(routing)
focuses on routing the demand onto the road network while accounting for
cyclist preferences and government priorities. Section
\@ref(road-segment-prioritisation) outlines a method for partitioning the
study area based on a community finding algorithm and routed cycling
demand. It then introduces the network expansion algorithms, and compares an approach grounded in
'egalitarianism' to one grounded in 'utilitarianism'.

@@

Many of the cycling network studies mentioned above use demand data for cycling as a starting point. 
Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely. 
@duthie2014optimization note that relying only on existing cycling activity to prioritise cycling
infrastructure can reinforce existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
entire study area. 
@olmos2020data opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match
the potential demand distribution to the distribution obtained from GPS data.

@@

For our purposes, we use a logistic regression model to calculate
potential cycling demand. The model was adapted from the Propensity to
Cycle Tool (PCT), which estimates the proportion of trips
($\boldsymbol{C_{p}}$) for each OD pair that would cycle under
different scenarios of change (we used the Government Target scenario,
representing a nationwide target of doubling cycling by 2025) as a
function of distance and hilliness [@lovelace2017propensity]. The
logistic regression model used to calculate $\boldsymbol{C_{p}}$ has
the following parameters:

```{=tex}
\begin{align}\label{eq:pcteqn}
     logit(C_{p}) = & \alpha + \beta_1 d + \beta_2 \sqrt{d} + \beta_3 d^2 + \beta_4 s + \beta_5 ds + \beta_6 \sqrt{d}s\nonumber 
\end{align}
```

\noindent where **d** and **s** are the distance and slope associated with each
OD pair and $\alpha$ and $\beta$ values are parameters calculated by a regression
model on training data. The square and square-root distance terms
"capture the non-linear impact of distance on the likelihood of
cycling", and interaction terms to capture the combined effect of slope
and distance [@lovelace2017propensity]. Alternative cycling uptake
models could be 'plugged in' to our approach for different contexts or scenarios of change.

@@

1.  **Cyclist Preference**: Work done by @dill2013four on examining
    cyclist typologies determined that around 60% of Portland residents
    fit under the *interested but concerned* category. These were people
    that enjoyed cycling but avoided it due safety concerns. The key to
    encouraging this group was to create a low-stress cycling network,
    not only though segregated infrastructure but also by planning
    routes that passed through residential streets.
2.  **Low-Traffic Neighborhoods**: The UK Department for Transport is
    allocating funding to local authorities to invest in Active
    Transport, partially through the creation of LTNs
    [@departmentgearchange2020]. This includes closing off residential
    streets to motorized traffic.
3.  **Existing Cycling Infrastructure**: Utilizing existing cycling
    infrastructure makes economic sense, as small investments may lead
    to large connectivity gains as the disconnected cycling
    infrastructure gets joined together.

The above points are accounted for by using a weighted road network for
routing. This has previously been done by multiplying all road segments
without cycling infrastructure by a fixed impedance factor
[@mauttone2017bicycle], or by assigning different weights to the road
segments proportional to the investment cost of bringing them to an
acceptable level of stress for cycling [@duthie2014optimization].
Perceived stress for cyclists has been found to be a function of vehicular traffic
volume and speed [@sorton1994bicycle], both of which vary predictably
with road type. @gehrke2020cycling build on this information to use
routing impedance factors for road segments that are based on road type
and the existence of cycling infrastructure.

For the purposes of this paper, we created a similar weighting profile
that is adjusted to favor less stressful roads (based on the definition
by @sorton1994bicycle and information from Table
\@ref(table:osmroadtypes)), and roads with existing cycling
infrastructure. We believe this to be more practical than the simplistic
approach adopted by @mauttone2017bicycle, as it goes beyond simply
favoring roads with existing cycling infrastructure enabling a hierarchy
of road preference based on perceived stress levels.

@@

\noindent A weighted distance $\boldsymbol{d_{w}}$ for each road
segment is calculated as follows ^[The **dodgr** r package
[@padgham2019dodgr] is used to route cycling demand onto the road
network. The package uses the OpenStreetMaps (OSM) road network and
allows the user to assign weights to roads based on their type. The
routing is done based on weighted shortest paths, with the distance
along each road segment being divided by a factor to obtain the weighted
distance for routing. It is more intuitive to multiply when weighting a
network, but the dodgr package divides by numbers between 0 and 1, which
achieves the same result. For the sake of reproducibility, we stick to
the convention used in the package.]:

\begin{equation}\label{eq:weight_distance}
    d_{w} = d_{unw}  \:/\:  W
\end{equation}

\noindent where $\boldsymbol{d_{unw}}$ is the unweighted distance and
$\boldsymbol{W}$ is the weight from Table \ref{table:weightprofiles}.

@@

```{r boxplotcircuity, fig.cap="Effect of Banning Cyclists from Trunk and Primary Roads for all OD Pairs (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/boxplot_weighted_unweighted_distances.png")
```

@@

It should be reiterated that the weighting profile used for routing has
been developed for the purposes of this study. It creates a hierarchy of
road preference that is grounded in cyclist preferences and government
plans to create LTNs. A wide range of weighting profiles could be used
to represent different types of cyclists and road environments, as
described in @gehrke2020cycling and @furth2016network. Sensitivity
analysis could be done to determine an optimal weighting profile, but
given the variation in city road
networks<!-- (Figure \ref{fig:perc_person-km}) -->, this would probably
require calibration to the specific city. More accurate routing could be
carried out given the availability of road-level data. In such cases we
would add additional impedance to specific roads, giving more useful
routing results than the current approach which considers all roads
of the same type to be equivalent.

@@

```{r communitiesmanchester, fig.cap="Communities Based on Potential Cycling Demand (Manchester)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/communities_alternative_Manchester.png")
```

@@

The results of the community detection are used to evaluate the
algorithms. This is done by looking at the *person-km satisfied* as
cycling infrastructure is added. Person-km is a measure of the total km
cycled on a road segment, so it is the product of the number of
potential commuters cycling on that road segment ($flow$) and the length
of the segment in km ($l$). For each road segment, the person-km is
equal to $flow * l$. In the case of Manchester, Table
\@ref(tab:personkmtable) shows that almost half of the person-km is in
community 1 , while only 0.5% of total person-km on the network is in
community 5.

@@

The approach can identify not only where there is high
potential for cycling but also trade-offs between
stress levels (related to motor traffic) and directness.
The results can therefore be used as a basis for recommendations on
road space reallocation *and* new infrastructure to unlock potential cycling demand.
The approach encourages consideration of a wider range of preferences
and needs than previous approaches that focus only on absolute potential.
Moreover, the inclusion of egalitarian principles in scenarios of change
encourages investment in cycling infrastructure to
increase the connectivity of existing cycling infrastructure
*and* investment that addresses geographical and social inequalities.
This ability to address inequalities in network prioritisation is particularly
important given research research showing substantial *inequalities*
around transport provision in general and cycling uptake and investment
in particular [@lucas2016method; @vidaltortosa_infrastructure_2020].

>>>>

---
title: Road Segment Prioritization for Bicycle Infrastructure
date: "`r Sys.Date()`"
geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. This paper demonstrates a new approach to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. Although each of these methods has been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach. The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios. A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
output:
  # bookdown::github_document2:
  bookdown::pdf_document2:
  # # bookdown::word_document2:
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
editor_options: 
  markdown: 
    wrap: 72
---

@@

The 2015 Paris agreement [@agreement2015paris] acknowledged that
fundamental changes to societies and economies are necessary to mitigate
climate change. Like other sectors, transport is under substantial
pressure to decarbonise, resulting in a number of technical innovations
including electric vehicles. But new vehicle technologies can only go so
far and do not tackle parallel problems such as congestion, road traffic
casualties and physical inactivity [@brand_climate_2020].

@@

> "Inadequate cycling infrastructure discourages cycling and wastes
 public money. Much cycling infrastructure in this country is
 inadequate. It reflects a belief, conscious or otherwise, that hardly
 anyone cycles, that cycling is unimportant and that cycles must take
 no meaningful space from more important road users, such as motor
 vehicles and pedestrians" [@departmentgearchange2020].

@@

Much research has been done to understand what gets people to cycle.
Segregated cycling infrastructure^[*Segregated cycling infrastructure*
refers to road space that is allocated to cyclists only, with physical
separation to protect cyclists from other modes of transport.] has been
shown to increase cycling uptake [@aldred2019impacts; @goodman2014new;
@marques2015infrastructure], with the separation from motorized
vehicles being key [@winters2011motivators]. Revealed preference of
cyclists shows that they are willing to deviate from the most efficient
routes in order to commute on safer roads [@crane2017longitudinal].
However, such deviations are only considered if they do not considerably
increase route circuitry; behaviour studies have found that the
probability of choosing a route decreases in proportion to its length
relative to the shortest route [@broach2011bicycle; @winters2010far].
Another defining feature for cycling infrastructure is how well
connected it is. Cyclists prefer cohesive infrastructure, particularly
when cycling on arterial roads with high levels of motorized traffic
[@stinson2003commuter]. The lack of well-connected cycling
infrastructure is one of the main obstacles to increasing cycling uptake
[@caulfield2012determining]. While direct and cohesive cycling networks
have been shown to positively impact cycling rates, density^[making an
area's bicycle network denser means adding more cycling routes in the
area and thereby giving cyclists more route options.] of the cycling
network is also vital [@schoner2014missing].

@@

The studies outlined above lay out the fundamentals for designing
cycling networks that generate significant cycling uptake, but they do
not propose network-level interventions. In this section we outline
methods used in past studies, namely optimization and network
analysis methods, such as connected components and community
detection, and examine how they are leveraged to suggest cycling network
designs. We compare the effectiveness of these network-level studies in
incorporating the fundamentals outlined above. 
Our proposed approach is inspired by these methods, but it attempts to add to them
by ensuring that all of the outlined fundamentals are accounted for. It
also goes further by attempting to factor in ethical considerations
related to the distribution of investment.

*Optimization* methods have been used to propose improvements to
cycling networks. @mesbah2012bilevel propose a bi-level formulation to
optimize allocation of cycling lanes to the network without exceeding a
set budget. They account for the effect of cycling lanes on car traffic,
and attempt to maximize utilization of said lanes with minimal impact on
car travel times. Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as "traffic evaporation" [@nello2020environmental].
On the other hand, @mauttone2017bicycle developed an optimization framework aiming at minimizing the total
user cost of cycling on the network without considering car usage. The aggregate flow^[*flow* is
used throughout this research to refer to the cycling demand when it is
routed onto the road network. The flow on any road segment is the
cumulative demand on it, resulting from cyclists commuting between
various Origin-Destination (OD) pairs.] on road segments (links) is obtained by using
shortest paths to route existing cycling demand onto the road network. The solution is a proposed set of links where cycling infrastructure
should be added in order to minimize the overall travel cost of cyclists
across the network. The cost of traversing a link is given as a function
of its length and whether or not it has cycling infrastructure. The
problem has also been solved by attempting to find the minimum cost of
improving roadway links to meet a desired level of service (LOS)
[@duthie2014optimization]. In this formulation, all origin-destination
(OD) pairs need to be connected by roads that meet the desired LOS. A
limitation of these approaches is that they do not explicitly solve for
continuity. The latter is addressed using either a constraint specifying
that each link with a bike lane should be connected to at least one
destination [@mesbah2012bilevel], a constraint on maximum deviation
from shortest paths [@duthie2014optimization], or a discontinuity
penalty to prioritize connected road segments [@mauttone2017bicycle].

In this paper, continuity is analysed by looking at the connectivity of
the network through the graph-theoretic concept of *connected
components*. @natera2019data for example,
study the existing cycling network in terms of its disconnected
components. They propose two different algorithms to connect disconnected
components by their most critical links and, in doing so, measure the
size of the growth of the largest connected component as a function of
the kilometers of network added. The concept of connected components is
also at the core of the methodology proposed by @olmos2020data. After
routing the cycling demand onto the network links, they use percolation
theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
filtering to identify the minimum flow at which the whole city is
connected by a giant component. While these approaches deal with
continuity better, they look at the network as a whole when attempting
to improve it, and in doing so fail to account for equitable
distribution of infrastructure.

@@

Our work builds on these *community finding* approaches by proposing a
similar greedy network expansion algorithm for cycle network expansion
within communities. We incorporate community finding methods for study
area partitioning with weighted routing to avoid links that are
stressful to cycle on. In doing so, we propose an approach that
accounts for motivators and deterrents to cycling. We propose three
sub-methods that address some of the limitations of previous
studies. These limitations include (a) an inherent bias of basing network
design solely on existing cycling demand, (b) the proposal of routes that may
not correspond to studies on cyclist preference and government policies,
and (c) an insufficient consideration of the ethical principles underlying
the analysis. The current work addresses these limitations through the following organisation. Section \@ref(calculating-potential-cycling-demand)
focuses on calculating potential cycling demand. Section \@ref(routing)
focuses on routing the demand onto the road network while accounting for
cyclist preferences and government priorities. Section
\@ref(road-segment-prioritisation) outlines a method for partitioning the
study area based on a community finding algorithm and routed cycling
demand. It then introduces the network expansion algorithms, and compares an approach grounded in
'egalitarianism' to one grounded in 'utilitarianism'.

@@

Many of the cycling network studies mentioned above use demand data for cycling as a starting point. 
Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely. 
@duthie2014optimization note that relying only on existing cycling activity to prioritise cycling
infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
entire study area. 
@olmos2020data opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match
the potential demand distribution to the distribution obtained from GPS data.

@@

For our purposes, we use a logistic regression model to calculate
potential cycling demand. The model was adapted from the Propensity to
Cycle Tool (PCT), which estimates the proportion of trips
($\boldsymbol{C_{p}}$) for each OD pair that would cycle under
different scenarios of change as a
function of distance and hilliness [@lovelace2017propensity]. We used the Government Target scenario,
indicating a nationwide target of doubling cycling by 2025. The
logistic regression model used to calculate $\boldsymbol{C_{p}}$ has
the following parameters:

```{=tex}
\begin{align}\label{eq:pcteqn}
     \logit(C_{p}) = & \alpha + \beta_1 d + \beta_2 \sqrt{d} + \beta_3 d^2 + \beta_4 s + \beta_5 ds + \beta_6 \sqrt{d}s 
\end{align}
```

\noindent where $d$ and $s$ are the distance and slope associated with each
OD pair, and $\alpha$ and all $\beta$s, are parameters calculated by a regression
model on the training data. The square and square-root distance terms
"capture the non-linear impact of distance on the likelihood of
cycling", and interaction terms to capture the combined effect of slope
and distance [@lovelace2017propensity]. Alternative cycling uptake
models could be 'plugged in' to our approach for different contexts or scenarios of change.

@@

1.  **Cyclist Preference**: Work done by @dill2013four on examining
    cyclist typologies determined that around 60% of Portland residents
    fit under the *interested but concerned* category. These were people
    that enjoyed cycling but avoided it due safety concerns. The key to
    encouraging this group was to create a low-stress cycling network,
    not only though segregated infrastructure but also by planning
    routes that passed through residential streets.
2.  **Low-Traffic Neighbourhoods (LTNs)**: The UK Department for Transport is
    allocating funding to local authorities to invest in Active
    Transport, partially through the creation of LTNs
    [@departmentgearchange2020]. This includes closing off residential
    streets to motorized traffic.
3.  **Existing Cycling Infrastructure**: Utilizing existing cycling
    infrastructure makes economic sense, as small investments may lead
    to large connectivity gains as the disconnected cycling
    infrastructure gets joined together.

The above points are accounted for by using a weighted road network for
routing. This has been previously done by multiplying all road segments
without cycling infrastructure by a fixed impedance factor
[@mauttone2017bicycle], or by assigning different weights to the road
segments proportional to the investment cost of bringing them to an
acceptable level of stress for cycling [@duthie2014optimization].
Perceived stress for cyclists has been found to be a function of vehicular traffic
volume and speed [@sorton1994bicycle], both of which vary predictably
with road type. @gehrke2020cycling build on this information to use
routing impedance factors for road segments that are based on road type
and the existence of cycling infrastructure.

For the purposes of this paper, we created a similar weighting profile
that is adjusted to favor less stressful roads (based on the definition
by @sorton1994bicycle and information from Table
\@ref(table:osmroadtypes)), and roads with existing cycling
infrastructure. We believe this to be more appropriate than the simplistic
approach adopted by @mauttone2017bicycle, as it makes use of a hierarchy
of road preference based on perceived stress levels, going beyond simply
favoring roads with existing cycling infrastructure.

@@

\noindent A weighted distance $d_{w}$ for each road
segment is calculated as follows ^[The **dodgr** r package
[@padgham2019dodgr] is used to route cycling demand onto the road
network. The package uses the OpenStreetMaps (OSM) road network and
allows the user to assign weights to roads based on their type. The
routing is done based on weighted shortest paths, with the distance
along each road segment being divided by a factor to obtain the weighted
distance for routing. It is more intuitive to multiply when weighting a
network, but the dodgr package divides by numbers between 0 and 1, which
achieves the same result. For the sake of reproducibility, we stick to
the convention used in the package.]:

\begin{equation}\label{eq:weight_distance}
    d_{w} = \frac{d_{unw}}{W}
\end{equation}

\noindent where $d_{unw}$ is the unweighted distance and
$W$ is the weight from Table \ref{table:weightprofiles}.

@@

```{r boxplotcircuity, fig.cap="Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/boxplot_weighted_unweighted_distances.png")
```

@@

It should be reiterated that the weighting profile used for routing has
been developed for the purposes of this study. It creates a hierarchy of
road preference that is grounded in cyclist preferences and government
plans to create LTNs. A wide range of weighting profiles could be used
to represent different types of cyclists and road environments, as
described in @gehrke2020cycling and @furth2016network. A sensitivity
analysis could be done to determine an optimal weighting profile, but
given the variation in city road
networks<!-- (Figure \ref{fig:perc_person-km}) -->, this would probably
require calibration to the specific city. More accurate routing could be
carried out given the availability of road-level data. In such cases we
would add additional impedance to specific roads, giving more useful
routing results than the current approach which considers all roads
of the same type to be equivalent.

@@

```{r communitiesmanchester, fig.cap="Communities based on potential cycling demand between MSOAs (Manchester)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
knitr::include_graphics("data/Manchester/Plots/communities_alternative_Manchester.png")
```

@@

The results of the community detection are used to evaluate the
algorithms. This is done by looking at the *person-km satisfied* as
cycling infrastructure is added. Person-km is a measure of the total km
cycled on a road segment, so it is the product of the number of
potential commuters cycling on that road segment ($flow$) and the length
of the segment in km ($l$). For each road segment, the person-km is
equal to $flow * l$. In the case of Manchester, Table
\@ref(tab:personkmtable) shows that almost half of the person-km is in
community 1, while only 0.5% of total person-km on the network is in
community 5.

@@

The approach can identify not only where there is high
potential for cycling but also trade-offs between
stress levels (related to motor traffic) and directness.
The results can therefore be used as a basis for recommendations on
road space reallocation *and* new infrastructure to unlock potential cycling demand.
The approach encourages consideration of a wider range of preferences
and needs than previous approaches that focus only on absolute potential.
Moreover, the inclusion of egalitarian principles in scenarios of change
encourages investment in cycling infrastructure to
increase the connectivity of existing cycling infrastructure
*and* investment that addresses geographical and social inequalities.
This ability to address inequalities in network prioritisation is particularly
important given research showing substantial *inequalities*
around transport provision in general and cycling uptake and investment
in particular [@lucas2016method; @vidaltortosa_infrastructure_2020].
